---
title: "R Markdown Analysis Template"
output: html_document
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, comment = NA)
```

## Setup

```{r}
library(tidyverse)
source("../R/ex_std_error.R")
```

Specify which data we will import and use:

```{r}
data_path <- "../data/raw/ex_mtcars.csv"
```

## Part 1: Tidy data

We'll do a short analysis using the `mtcars` dataset. After loading the data, we'll convert `vs` and `am` (which encode engine and transmission types) to factors, since they are really dichotomous variables, not numbers.

R Markdown allows you to write code in a format that is meant to be read. To make it as easy as possible to follow, use a predictable format:

- Introduce the process with a function that composes each step (`tidy_mtcars_pipe()`)
- Write each step as a pure function of the full dataset, in the order they are called
- End by loading the data and running the pipeline

That way you present the reader with an overview at the beginning, fully explain the steps in the middle, and restrict all side effects (file I/O, assigning global variables) to the end.

```{r}
tidy_mtcars_pipe <- function(df) {
  df %>% 
    recode_vs() %>% 
    recode_am()
}
```

```{r}
recode_vs <- function(df) {
  df %>% 
    mutate(vs = factor(vs)) %>% 
    mutate(vs = fct_recode(vs, `V` = "0", `straight` = "1"))
}

recode_am <- function(df) {
  df %>%
    mutate(am = factor(am)) %>% 
    mutate(am = fct_recode(am, `auto` = "0", `man` = "1"))
}
```

Run the pipeline:

```{r}
ex_mtcars <- read_csv(data_path)

better_mtcars <-
  ex_mtcars %>% 
  tidy_mtcars_pipe()

better_mtcars %>% 
  saveRDS("../data/working/ex_better_mtcars.rds")
```

## Part 2: Summary statistics and plots

Now we investigate the relationship between engine type `vs` and horsepower `hp`.

We calculate mean and standard error, then plot them:

```{r}
plot_pipe <- function(df) {
  df %>% 
    vs_hp_mean_se() %>% 
    vs_hp_plot()
}
```

```{r}
# `ex_std_error()` is a helper function in another script
vs_hp_mean_se <- function(df) {
  df %>% 
    group_by(vs) %>% 
    summarize(mean_hp = mean(hp), std_err_hp = ex_std_error(hp))
}

# a boring and poorly-designed plot
vs_hp_plot <- function(df) {
  df %>% 
    ggplot(aes(
      x = vs,
      y = mean_hp,
      ymin = mean_hp-(1.96*std_err_hp),
      ymax = mean_hp+(1.96*std_err_hp)
    )) +
    geom_col() +
    geom_errorbar(width = 0.3)
}
```

Plot the association between engine type and horsepower:

```{r}
better_mtcars %>% 
  plot_pipe()
```

Save the summary dataset as example output:

```{r}
better_mtcars %>% 
  vs_hp_mean_se() %>% 
  saveRDS("../output/ex_engine_hp.rds")
```

## Testing

Testing is easy with `testthat::test_file`:

```{r}
library(testthat)

test_file("../tests/ex_test_mtcars.R")
```

Or `assertr` for nice data verification options:

```{r}
library(assertr)

better_mtcars %>% 
  verify(is.logical(vs), error_fun = error_append) %>% # vs is a factor
  verify(nrow(.) == 31, error_fun = error_return)      # mtcars has 32 rows
```

